visitation_bias

### 2021-08-25 code/ms2_nullmodels.Rmd
preped ms for Ecol Lett

view(filter(vis_spp, ktype == "1"))

ggplot(filter(vis_spp, ))

# For setting up the HPC
# 
# bb_traits <- read_csv("./bb_traits.csv") %>%
#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
# 
# site_data <- read_csv("./site_data.csv") %>%
#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
#   mutate(site = factor(site),
#          tree_line = factor(tree_line))
```

@@

ggsave("../ms_2/figures/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)

# Species-wise (3-col)
ggplot(null_tab1, aes(Observed, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~fct_rev(bb.sp), ncol = 3) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_spwise_preference_3col.pdf", height = 6, width = 5)

## Morphotype-wise
ggplot(null_tab1, aes(Observed, bb.sp)) +
  geom_segment(aes(y = bb.sp, yend = bb.sp, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~ktype, ncol = 4) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)

@@

## 2.2 Overall bias by morphotype
```{r}
null_model_overall<- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Selectivity")

@@

# Plot
null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Selectivity", 
       color = "Species", linetype = "p < 0.05") +
  theme_light(10)

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)

@@

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, date, bb.sp, ktype, Observed, Null, Lower.95.CL, Upper.95.CL, Test, 
         SES, preference, preference.upper95, preference.lower95) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, date, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>% #There are a few cases in which total.observed does not equal unity. This is because bees occasionally visited plants in ktypes 0, 9, or 10, which we dropped from the data prior to null model calculation but after calculating proportional visitation. This is as it should be.
  ungroup() %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation; I checked, and every time Null = 0, Observed = 0 also. 
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         bb.ktype = factor(bb.ktype),
         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()

ggplot(null_tab3, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")
```

@@

ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
```

@@

raw_visits_overall <- ggplot(visitation_prop, aes(ktype, mean.prop.visits)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
  geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +
  theme_light(10) +
  labs(x = "Floral morphotype",
       y = "Mean proportional visitation")

@@

ggsave("../ms_2/figures/overall_ktype.pdf", overall_ktype, height = 4.5, width = 9.5)
ggsave("../ms_2/figures/overall_ktype.png", overall_ktype, height = 4.5, width = 9.5)
```

@@

# Plot
raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species", linetype = "p < 0.05") +
  theme_light(10)

@@

# Binomial model
gam6 <- bam(prop.cover ~
              ktype +
              s(site, by = ktype, bs = "re") +
              s(year, by = ktype, bs = "re") +
              te(yday, elev.mean, by = ktype),
            #family = betar(eps = 0.001),
            family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(flower_cover_ktype.prop, 
                            ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam6)
gam6_sum <- summary(gam6, re.test = FALSE) 
plot.gamViz(gam6)

@@

ggsave("../ms_2/figures/flower_cover_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/flower_cover_gam.png", height = 5, width = 6.5)
```

>>>>

fl_tax <- read_csv("../data/processed_data/floral_tax.csv")
```

@@

ggsave("../ms_2/figures/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)
# 
# # Species-wise (3-col)
# ggplot(null_tab1, aes(Observed, ktype)) +
#   geom_segment(aes(y = ktype, yend = ktype, 
#                    x = Lower.99.CL, xend = Upper.99.CL), 
#                color = "gray40", inherit.aes = FALSE) +
#   geom_point(aes(color = Test)) +
#   facet_wrap(~fct_rev(bb.sp), ncol = 3) +
#   scale_y_discrete(limits = rev) +
#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
#   labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
#   theme_light()
# 
# ggsave("../ms_2/figures/null_model_spwise_preference_3col.pdf", height = 6, width = 5)
# 
# ## Morphotype-wise
# ggplot(null_tab1, aes(Observed, bb.sp)) +
#   geom_segment(aes(y = bb.sp, yend = bb.sp, 
#                    x = Lower.99.CL, xend = Upper.99.CL), 
#                color = "gray40", inherit.aes = FALSE) +
#   geom_point(aes(color = Test)) +
#   facet_wrap(~ktype, ncol = 4) +
#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
#   labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
#   theme_light()
# 
# ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
# ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)

@@

## 2.2 Overall bias by morphotype
```{r}
null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(12) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Preference")

@@

# Plot
null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Preference", 
       color = "Species", linetype = "p < 0.05") +
  scale_x_continuous(breaks = c(7, 10, 13, 16)) +
  theme_light(12)

ggsave("../ms_2/figures/fig3.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)

@@

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, date, bb.sp, ktype, Observed, Null, Lower.95.CL, Upper.95.CL, Test, 
         SES, preference, preference.upper95, preference.lower95) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, date, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>% #There are a few cases in which total.observed does not equal unity. This is because bees occasionally visited plants in ktypes 0, 9, or 10, which we dropped from the data prior to null model calculation but after calculating proportional visitation. This is as it should be.
  ungroup() %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation; I checked, and every time Null = 0, Observed = 0 also. 
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         bb.ktype = factor(bb.ktype),
         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

@@

ggsave("../ms_2/figures/fig5.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
```

@@

raw_visits_overall <- ggplot(visitation_prop, aes(ktype, mean.prop.visits)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
  geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +
  theme_light(12) +
  labs(x = "Floral morphotype",
       y = "Mean proportional visitation")

@@

ggsave("../ms_2/figures/fig2.pdf", overall_ktype, height = 4.5, width = 9.5)
ggsave("../ms_2/figures/overall_ktype.png", overall_ktype, height = 4.5, width = 9.5)
```

@@

# Plot
raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species", linetype = "p < 0.05") +
  scale_x_continuous(breaks = c(7, 10, 13, 16)) +
  theme_light(12)

@@

# Binomial model
# gam6 <- bam(prop.cover ~
#               ktype +
#               s(site, by = ktype, bs = "re") +
#               s(year, by = ktype, bs = "re") +
#               te(yday, elev.mean, by = ktype),
#             #family = betar(eps = 0.001),
#             family = "binomial",
#             method = "fREML",
#             discrete = TRUE,
#             data = filter(flower_cover_ktype.prop, 
#                             ktype %in% c("3", "5", "6", "7"))) %>% 
#   getViz()
# 
# check.gamViz(gam6)
# gam6_sum <- summary(gam6, re.test = FALSE) 
# plot.gamViz(gam6)

@@

ggsave("../ms_2/figures/fig6.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/flower_cover_gam.png", height = 5, width = 6.5)
```
